### 2019-03-23 micro4_2.tex
Micro

\begin{center}
	Microeconomics 4. Problem Set 1.
\end{center}
\section*{1 The Doctor and the Patient}
Consider the cheap talk signalling model a-la Crawford and Sobel (1982). Suppose that
there is a patient that does not feel well and goes to see a doctor. The doctor should
decide which treatment $a$ the patient needs. The type of treatment will depend on
the health state of the patient $\theta$, which is only privately observed by the patient. The
doctor's prior about the patient's health is $\theta \sim U[0; 1]$. The patient can send a costless,
non-binding and non-verifiable signal $s \in [0; 1]$ about his health state (i.e. explain how
he feels). The doctor cares about her reputation, so she wants to prescribe a appropriate
treatment for the patient's health: $U
D(a; \theta) = -(a-\theta)^2$. Instead, the patient, that is
very hypochondriac, wants to be overtreated: $U^P(a; \theta) = -(a - (\theta + b))^2$, where $b > 0$ is
the bias towards excessive medication.
\begin{enumerate}
	\item Assume that b is such that there exist a 2-partitional equilibrium $(p = 2)$. Characterize this equilibrium.
	\item Under which condition will such an equilibrium exist?
	\item Is this the only equilibrium that exists? If yes, argue why. If no, characterize the
	other(s) equilibrium (equilibria).
	\item Compute the expected payoff of the doctor and the patient. How do the payoffs
	change with $b$?
\end{enumerate}

@@

	Since outside options of workers are 0, the optimal contracts monopoly proposes to workers should be such that

@@

	That means that at any interior solution high type has incentive to deviate and mimic to low type. Hence we should add incentive compatibility constraints. The IC constraints are:

@@

	c_1(x_1^*) - c_2(x_1^*) \le 0\\
	c_2(x_2^*) - c_1(x_2^*) \le 0

@@

	Since $c_2(x) > c_1(x)\ \forall\ x > 0$ ($g(x) = c_2(x) - c_1(x)$, $g(0) = 0, g'(x) = c'_2(x) - c_1'(x) > 0$ hence $g(x)$ is strictly increasing, since $g(0) = 0$ then $g(x)>0\ \forall\ x > 0$) it follows that $x_2^* = 0$. That means that monopoly should offer the following contracts:

@@

	&(0, 0)\ \text{ which is supposed to be for low type}\\
	&(w_1 = c_1(c_1'^{-1}(1)), x_1 = c_1'^{-1}(1))\ \text{ which is supposed to be for high type}

@@

@@

@@

		\frac{x_1^2}{2} - x_1^2 \le 0\\
		x_2^2 - \frac{x_2^2}{2} \le 0

@@

		x_2^* = 0, x_1^* = 1\\
		w_2 = 0, w_1 = \frac{1}{2}

@@

\item If high type is more risk averse, then for a given wealth level in good state he will demand more wealth in bad state comparing with utility $u(W)$ to maintain the same utility level. That means that new separating equilibrium looks like as follows:

@@

	\section*{The Patient}
		Suppose that the hypochondriac patient of question 1 is a professional runner. The
		patient has been training for a very important race for the last two years and he is very
		much looking forward to it. In order to run in this important race, first he should be
		cleared that he is "fit for running" by the Organization of the race. Suppose that when
		the patient meets his doctor, there is also a member of the Organization in the surgery.
		What do you think will happen to the quality of the information transmitted by the
		patient?

@@

		Since the patient is very much looking forward to taking part in the race, he does not want to be declared unfit for running. That means parameter $b$ becomes smaller. Because if patient complains about his health a lot, the member of the Organization will likely to declare him as unfit for running. And the sportsman understands it. Thus, $b$ becomes smaller and as we can see in Problem 1, expected profits increases, i.e. the quality of information transmitted goes up.

@@

			We should use some refinements i.e. slightly alter the concept of equilibrium. We can use anticipatory (Wilson, JET, 1977 ) and reactive (Riley, Econometrica, 1979 ) equilbrium concepts. The anticipatory equilibrium concept manages to deal with absence of pooling equilibria. Once the previous contracts are withdrawn under this concept, firm skim off not only good type but also bad type, but in this situation firms begin losing money, so the deviation is no longer profitable. That is, once contracts are withdrawn firms cannot skim off only a good type.

			The concept of reactive equilibrium manages to deal with the second problem. This concept allows firms to add more contracts. Then, if we are in the case when separating equilibrium is beaten by pooling, adding another contract (see Fig. \eqref{fig5}, point 2) makes the initial deviation no longer profitable. Hence the separating equilibrium is not beaten by pooling under this concept.
				\begin{figure}[H]
				\centering
				\includegraphics[width=0.8\textwidth]{plotdraft}
				\caption{}\label{fig5}
			\end{figure}
\end{document}

>>>>

\begin{center}
	Microeconomics 4. Problem Set 1.
\end{center}
\section*{1 The Doctor and the Patient}
Consider the cheap talk signalling model a-la Crawford and Sobel (1982). Suppose that
there is a patient that does not feel well and goes to see a doctor. The doctor should
decide which treatment $a$ the patient needs. The type of treatment will depend on
the health state of the patient $\theta$, which is only privately observed by the patient. The
doctor's prior about the patient's health is $\theta \sim U[0; 1]$. The patient can send a costless,
non-binding and non-verifiable signal $s \in [0; 1]$ about his health state (i.e. explain how
he feels). The doctor cares about her reputation, so she wants to prescribe a appropriate
treatment for the patient's health: $U^D(a; \theta) = -(a-\theta)^2$. Instead, the patient, that is
very hypochondriac, wants to be overtreated: $U^P(a; \theta) = -(a - (\theta + b))^2$, where $b > 0$ is
the bias towards excessive medication.
\begin{enumerate}
	\item Assume that b is such that there exist a 2-partitional equilibrium $(p = 2)$. Characterize this equilibrium.
	\item Under which condition will such an equilibrium exist?
	\item Is this the only equilibrium that exists? If yes, argue why. If no, characterize the
	other(s) equilibrium (equilibria).
	\item Compute the expected payoff of the doctor and the patient. How do the payoffs
	change with $b$?
\end{enumerate}

@@

	Since outside options of workers are 0, the optimal contracts monopoly proposes to workers in case of complete information should be such that

@@

	That means that at any interior solution high type has incentive to deviate and mimic to low type. Hence we should add incentive compatibility and participation constraints.

@@

	w_1^* - c_1(x_1^*) \ge 0\\
	w_2^* - c_2(x_2^*) \ge 0

@@

	Since $c_2(x) > c_1(x)\ \forall\ x > 0$ ($g(x) = c_2(x) - c_1(x)$, $g(0) = 0, g'(x) = c'_2(x) - c_1'(x) > 0$ hence $g(x)$ is strictly increasing, since $g(0) = 0$ then $g(x)>0\ \forall\ x > 0$) it follows that $w_1^* - c_1(x_1^*) \ge w_2^* - c_1(x_2^*) > w_2^* - c_2(x_2^*) \ge 0$. Thus, $w_1^*-c_1(x_1^*) > 0$ is implied by other constraints. Moreover, if $w_2^* - c_2(x_2^*) > 0$ then one can slightly decrease both $w_1^*, w_2^*$ such that all constraints hold but the objective function increases. Hence $w_2^* = c_2(x_2^*)$. On top of that, if $w_1^* - c_1(x_1^*) > w_2^* - c_1(x_2^*)$ then we can slightly decrease $w_1^*$ such that all constraints hold but the objective function increases. That means that $w_1^* - c_1(x_1^*) = w_2 - c_1(x_2^*)$. Finally, the monopoly optimization problem looks as follows:

@@

	\underset{(x_1, w_1), (x_2, w_2)}{\max}\ \lambda(x_1 - w_1) + (1 - \lambda) (x_2 - w_2)\\
	w_1 - c_1(x_1) = c_2(x_2) - c_1(x_2)\\
	w_1 - c_2(x_1) \le 0\\
	c_2(x_2) = w_2

@@

	FOCs imply:
	c'_1(x_1^*) = 1\\
	\lambda c_1'(x_2^*) + 1 - \lambda - c_2'(x_2^*) = 0
	If functions $c_1(\cdot), c_2(\cdot)$ are given it is easy to find optimal contracts from FOCs.

@@

		\underset{x_1, w_1, x_2, w_2}{\max}\ \frac{1}{2}\left(x_1 -w_1\right) + \frac{1}{2} \left(x_2 - x_2^2\right)\\

@@

		w_1 - \frac{x_1^2}{2} = \frac{x_2^2}{2}\\
		w_1 - x_1^2 \le 0\\
		w_2 = x_2^2

@@

		x_2^* = \frac{1}{3}, x_1^* = 1\\
		w_2 = \frac{1}{9}, w_1 = \frac{5}{9}

@@

\item If high type is more risk averse, then for a given wealth level in good state he will demand more wealth in bad state comparing with utility $u(W)$ to maintain the same utility level. Put it another way, if less risk averse consumer is indifferent between full insurance and some other point, then more risk averse consumer strictly prefers complete insurance. That means that new separating equilibrium looks like as follows:

@@

	\section*{5 The Patient}
		Suppose that the hypochondriac patient of question 1 is a professional runner. The
		patient has been training for a very important race for the last two years and he is very
		much looking forward to it. In order to run in this important race, first he should be
		cleared that he is "fit for running" by the Organization of the race. Suppose that when
		the patient meets his doctor, there is also a member of the Organization in the surgery.
		What do you think will happen to the quality of the information transmitted by the
		patient?

@@

		Since the patient is very much looking forward to taking part in the race, he does not want to be declared unfit for running. That means parameter $b$ effectively becomes smaller (of course parameter $b$ is the innate trait of the agent and cannot change, but once the member of Organization is in the surgery, the patient becomes to obtain disutility from the overtreating and declaring unfit for the race, i.e. his utility becomes $U^P = - (a - (b + \theta - m))^2$ where $m > 0$. If we denote $b' = b - m$ then his effective bias $b'$ becomes smaller once the member of Organization is in surgery). Because if patient complains about his health a lot, the member of the Organization will likely to declare him as unfit for running. And the sportsman understands it. As is has been shown in Problem 1, once bias becomes smaller, expected profits increases. The same result was shown for a more general set-up by Crawford and Sobel (1982):
		\begin{quote}
			There may be a good case for presuming that direct communication is more likely to play an important role, the more closely related are agents' goals.
		\end{quote}
	That is, the quality of information transmission is likely to become higher.

@@

			We should use some refinements i.e. slightly alter the concept of equilibrium. We can use anticipatory (or E2) (Wilson, JET, 1977 ) and reactive (Riley, Econometrica, 1979) equilibrium concepts.
			Wilson (1977) defines the anticipatory equilibrium as follows:
			\begin{quote}
				Each policy in the equilibrium set must earn nonnegative profits and there must not be a new set of policies which will earn nonnegative profits for each policy and strictly positive profits for some new policy after the policies in the original set which have become unprofitable are withdrawn
			\end{quote} 
		As a result, this concept manages to deal with absence of pooling equilibria. Once the previous contracts are withdrawn under this concept, firm skim off not only good type but also bad type, but in this situation firms begin losing money, so the deviation is no longer profitable. That is, once contracts are withdrawn firms cannot skim off only a good type.
			Riley (1979) defines the concept of reactive equilibrium as follows:
			\begin{quote}
				A set of offers is a reactive equilibrium if, for any additional offer which generates as expected gain to the agent making the offer, there is another which yields a gain to a second agent and losses to the first. Moreover, no further addition to the set of offers generates losses to the second agent.
			\end{quote}
			This concept manages to deal with the second problem. It allows firms to add more contracts. Then, if we are in the case when separating equilibrium is beaten by pooling, adding another contract (see Fig. \eqref{fig5}, point 2) makes the initial deviation no longer profitable. Hence the separating equilibrium is not beaten by pooling under this concept.
				\begin{figure}[H]
				\centering
				\includegraphics[width=0.8\textwidth]{plotdraft}
				\caption{}\label{fig5}
			\end{figure}

		\section*{7 Single-crossing}
			Most screening models assume single-crossing condition. Think about what this condition
			would require in real life. Do you think single-crossing is always a reasonable assumption?
			Explain why it is reasonable in any application, or suggest a real-life example where this assumption might fail and explain why.

			This condition requires that whatever effort people have already put to signal of their type, for putting the additional amount of effort the low type always requires more compensation than the high type. I think that it is a reasonable assumption in real-life terms but only starting from some point. Cost of some signals can be very low, as a result, both groups of people will demand the same amount of compensation for completing these tasks. For example, almost everybody graduates from school and the vast majority then enter in some universities. The costs of this signals are very low, so people of both types of abilities bear the similar costs of signaling. But starting from some point (on effort line), tasks become more complicated and require more time and cognitive abilities, in this case people of high type will spend less time, put less effort to solve such a kind of tasks. For example just to graduate from school is a very easy task, however graduate from school with some kind of honors (gold medal or smth like that) is not so easy, as a result the majority of people who succeed in it are either of high type or of low type but who put a lot of effort (maybe these efforts transformed them to people of high type) and here the single crossing condition is satisfied. I think that true indifference curves should look something like this:
			 \begin{figure}[H]
			 	\centering
			 	\includegraphics[width=0.8\textwidth]{plotdraft}
			 	\caption{}\label{fig6}
			 \end{figure}
\end{document}
